Topological Graph Theory

### 2018-03-20 Algebra/Homework/Homework.tex
what da hell

Let $a,b\in\mathbb R$ be given. Now let us investigate the derivative of $x^2 + ax + b$ with respect to $x$. Any student of calculus can tell you $\derivative{x^2+ax+b}{x} = 2x+a$. Now consider that when $x < \frac a2$

\subsubsection{Question 23}
\subsubsection{Question 28}
\subsubsection{Question 29}

>>>>

Let $f(x) = x^2+ax+b$, thus $f'(x) = 2x+a$. $f'(x)$ is linear so there exists only one $x\in\mathbb R$ for which $f'(x) = 0$, and thus this $x$ is a global extrema for $f$, so $f$ can not be surjective. Now consider $x_1 = -\frac a2-1$ and $x_2=-\frac a2+1$, thus
\begin{align*}
f(x_1) &= \left(-\frac a2-1\right)^2+a\left(-\frac a2-1\right)+b \\
&= \frac{a^2}4+2\frac a2+1-\frac{a^2}2-a+b\\
&=\frac{a^2}4+1+b\\
f(x_2) &= \left(-\frac a2+1\right)^2+a\left(-\frac a2+1\right)+b \\
&= \frac{a^2}4-2\frac a2+1-\frac{a^2}2+a+b\\
&= \frac{a^2}4+1+b
\end{align*}

so $f$ must be 1-1.

\subsubsection{Question 23}

\def\RB{}
\def\LB{}

\marginnote{Ugly proof:}
First let us show that there exists some bijection from $\mathbb N$ to $\ZNonNegative^2$. Consider the 1 norm on $\ZNonNegative^2$, defined as $\norm{(a,b)}_1 = a + b$. Then we may partition $\ZNonNegative^2$ into subsets $P_n = \setbuilder{x\in\ZNonNegative^2}{\norm x_1 = n}$, for any $n\in \ZNonNegative$. Notice that for $(a,b)\in P_n$ then $a \le n$ and $b \le n$, thus forcing $P_n$ to be finite. Now we can construct a function mapping from $\mathbb N$ to $\ZNonNegative^2$ by giving each element of $P_1$ a number from $1$ to $\card{P_0}$ (inclusive), then the next $\card{P_1}$ will be given to elements of $P_1$ and so on infinitely. Notice that by construction $x\not= y \implies f(x) \not= f(y)$, so we get this being 1-1, additionally for any $(a,b) \in \ZNonNegative^2$, $(a,b)\in P_{a+b}$ and thus receives a number greater than $\sumfrom{n=0}{a+b-1}{\card{P_n}}$ and less than or equal to $\sumfrom{n=0}{a+b}{\card{P_n}}$. This means that we can label each element of $\ZNonNegative$ with a single natural number and thus have a bijection.

Now we can also construct a trivial bijection, $h:\ZNonNegative^2\to S$ as $h(a,b) = 2^a3^b$. Now we may compose the bijections to get a 1-1 correspondence $\mathbb N = S$ onto $T$.

\bigskip

\marginnote{Nice proof:}
First notice that $T \subset S$ so there exists the trivial injective function from $T$ to $S$. Second notice that $f:S\to T$ defined as $f(s)=2^s$ is both well defined as injective. By the  \href{https://en.wikipedia.org/wiki/Schr\%C3\%B6der\%E2\%80\%93Bernstein_theorem}{Schr\"oder-Bernstein theorem} there must be some bijection from $S$ to $T$.

\subsubsection{Question 28}

Let $S$ be a finite set, with $f:S\to S$. Now let $f(x) = f(y)$, for some $x\not=y$, then there remain $\card S - 2$ elements in $S\setminus\set{x,y}$ and $\card S - 1$ elements in $S\setminus\set{f(x)}$. This means that for any definition of $f$ on $S\setminus\set{x,y}$ it can not possibly be onto $S\setminus\set{f(x)}$. We have now shown $f$ not being 1-1 implies $f$ not being onto, by contrapositive $f$ being onto implies $f$ is 1-1.

\subsubsection{Question 29}

Let $S$ be a finite set, with $f:S\to S$ injective. Now as $f$ is 1-1 each $s\in S$ has a unique $f(s)\in S$, so $f(S)$ must have exactly $\card S$ unique elements, thus $f(S) \subset S$ with exactly $\card S$ elements.\footnote{$f(A)$ is defined as $\setbuilder{y\in\range{f}}{\Exists{x\in \domain{f}}{f(x) = y}}$ when $A \subset \domain{f}$ and $A\not\in\domain{f}$.} Because $S$ is finite, this implies $f(S) = S$.

\subsection{Section 4}

\subsubsection{Question 5}

\question{(a)} First identity:
\begin{align*}
f^2g^2&=ffgg\\
&=f(fg)g\\
&=f(gf)f\\
&=(fg)^2
\end{align*}

\question{(b)} Second Identity: Let $i$ be the identity function.
\begin{align*}
\inv f\inv ggf &= i\\
\inv f\inv ggf\inv{(gf)} &= i\inv{(gf)}\\
=\inv f\inv g i&=i\inv{(fg)}\\
=\inv f\inv g&=\inv{(fg)}
\end{align*}

\subsubsection{Question 9}
\subsubsection{Question 10}
\subsubsection{Question 14}
\subsubsection{Question 21}
\subsubsection{Question 27}
\subsubsection{Question 30}
\subsubsection{Question 32}

### 2018-03-20 Topology/Paper/glossary.tex
what da hell

\newglossaryentry{set}
{
	name={set},
	description={A collection of objects}
}
\newglossaryentry{vertex}
{
	name={vertex},
	description={A point or node in a graph},
	plural={verticies}
}
\newglossaryentry{edge}
{
	name={edge},
	description={A connection or line between points in a graph}
}
\newglossaryentry{connected}
{
	name={connected},
	description={A graph where one may start at any point and could follow edges and eventually get to any other point}
}
\newglossaryentry{multigraph}
{
	name={multigraph},
	description={Like a graph, but multiple \glspl{edge} may connect the same \gls{vertex} pair, and an edge may connect a \glx{vertex} to itself}
}
\newglossaryentry{complete}
{
	name={complete graph},
	description={A graph with all possible \glspl{edge} included, the notation $k_n$ is used to denote the \gls{complete} with $n$ \glspl{vertex}}
}
\newglossaryentry{face}
{
	name={face},
	description={An section of the space we are embedding our graph in that is separated from the rest of the space by \glspl{edge}}
}
\newglossaryentry{contraction}
{
	name={contraction},
	description={A graph operation where one removes an edge by fusing two \glspl{vertex} together}
}
\newglossaryentry{subgraph}
{
	name={subgraph},
	description={$A$ is a \gls{subgraph} of $B$ if $A$'s \gls{vertex} \gls{set} and \gls{edge} \gls{set} are subsets of $B$'s \gls{vertex} \gls{set} and \gls{edge} \gls{set} respectively}
}

>>>>

\newglossaryentry{set}
{
	name={set},
	description={A collection of objects}
}
\newglossaryentry{vertex}
{
	name={vertex},
	description={A point or node in a graph},
	plural={verticies}
}
\newglossaryentry{edge}
{
	name={edge},
	description={A connection or line between verticies in a graph}
}
\newglossaryentry{connected}
{
	name={connected},
	description={A graph where one may start at any point and could follow edges and eventually get to any other point}
}
\newglossaryentry{multigraph}
{
	name={multigraph},
	description={Like a graph, but multiple edges may connect the same vertex pair, and an edge may connect a vertex to itself}
}
\newglossaryentry{complete}
{
	name={complete graph},
	description={A graph with all possible edges included, the notation $K_n$ is used to denote the complete with $n$ verticies}
}
\newglossaryentry{face}
{
	name={face},
	description={An section of the space we are embedding our graph in that is separated from the rest of the space by edges}
}
\newglossaryentry{contraction}
{
	name={contraction},
	description={A graph operation where one removes an edge by fusing two verticies together}
}
\newglossaryentry{subgraph}
{
	name={subgraph},
	description={$A$ is a subgraph of some graph $\mathcal G$ iff you could add verticies and edges to $A$ and somehow get $\mathcal G$}
}
\newglossaryentry{graph}
{
	name={graph},
	description={A set of verticies and edges}
}

### 2018-03-20 Topology/Paper/paper.tex
what da hell

\documentclass{article}
\input{../../definitions}
\input{glossary}
\title{Topological Graph Theory}
\begin{document}
\maketitle
\tableofcontents

@@

\section{Overview}
\subsection{Graphs}
Before we talk about topological graph theory with any level of understanding we must first understand what a graph is.

A graph is generally defined as a \gls{set} of \glspl{vertex} combined with a \gls{set} of \glspl{edge} between \glspl{vertex}, however here it may be more useful to think about them visually with a simple representation.

@@

Once we have drawn this we have a representation of a graph. If we were to move the \glspl{vertex} around on the paper but leave them having the same \glspl{edge} (the same \glspl{vertex} are connected to the point as they were before). we would be left with the same graph. That is to say, it doesn't mater where we put a \gls{vertex} on our sheet, the graph exists independently of the representation we draw.

@@

I first came across this question in the 8\textsuperscript{th} grade, and it was presented to us during geometry class. While undoubtedly an interesting problem, it is quite misleading to try and think of this as a geometry problem. Instead we will try and reduce it to a graph problem.

Let us start by thinking of every island as a \gls{vertex} and every bridge as an \gls{edge}. We find the following graph.

@@

It is worth noting two things about the above diagram. First that the labels on the \glspl{vertex} and \glspl{edge} are unrelated to the problem, but have been added simply to make referring to parts of the graph much easier. Second that whatever the above image depicts, does not fit our definition of a graph. %TODO add glossary entry for graph.

Notice that \gls{edge} $2$ and \gls{edge} $3$ both connect \gls{vertex} $A$ to $B$, as well \glspl{edge} $5$ and $6$ do for $B$ and $C$. This is a strict violation of our definition for a graph. The issue of course then comes to what would one call such a beast as this where, presumably, one is able to have as many connections between any pair of \glspl{vertex} and could even have connections from a \gls{vertex} to itself.

I am particularly glad that you're paying enough attention to notice that the diagram does not depict a graph. This is what we will refer to as a \gls{multigraph}. It is worth noting that graphs are a type of \gls{multigraph}, so anything we show to be true for all \glspl{multigraph}, is also true for all graphs.

@@

Consider the following graph.

@@

This graph is called a \gls{complete} as every \gls{vertex} is connected to every other \gls{vertex} by an \gls{edge}; in fact this graph in particular is called $K_4$ as it is the complete four vertex graph. We would like to find out if we can draw this above graph without having any lines crossing. We can in fact draw this graph without any intersections and for any skeptics who may being reading this, the below is $k_4$ without any \glspl{edge} intersecting.

@@

So if this graph can be drawn without intersection, can any graph be drawn without intersections? If some can and some can't how do we tell which can be drawn and which can not?

\subsection{Kuratowski's Theorem}
Kuratowski's theorem tell us exactly what graphs can be drawn on a plane or sphere. We say either a plane or sphere as it turns out to be the same question. %TODO prove that drawing on plane and sphere is same question
Now before we tackle the theorem let's lay down some terminology.

@@

We say this graph has two \glspl{face}. One \gls{face} is easy to see and contained within triangle $ABC$. The other \gls{face} is a bit harder to see, however it is all the space outside of the triangle $ABC$. We may call this second face the outer \gls{face} for this drawing of the graph, however it is also possible to draw it such that the inner face may be the outer face. You can think of this as kind of turning the triangle inside out.

A face only makes sense to talk about for a graph that can be embedded in a certain space. For example the graph $K_4$ may look like it has $5$ faces in the following drawing.

@@

	\path[every node/.style={font=\sffamily\small}]
	(1) edge node {} (4)
	edge node {} (3)
	(2) edge node {} (1)
	edge node {} (4)
	(3) edge node {} (2)
	(4) edge node {} (3);
	\end{tikzpicture}
\end{center}
This would be wrong as a face assumes a drawing without \glspl{edge} intersecting. This can be rectified if the graph can be drawn without intersection and thus we find this drawing.
\begin{center}
	\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
	thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]

@@

	\path[every node/.style={font=\sffamily\small}]
	(1) edge node {} (4)
	edge [bend right = 100]node {} (3)
	(2) edge node {} (1)
	edge node {} (4)
	(3) edge node {} (2)
	(4) edge node {} (3);
	\end{tikzpicture}
\end{center}
This means there are actually $4$ \glspl{face} in $k_4$.

@@

Let us consider the following graph.

@@

This can then be reduced into a graph again by simply treating $b_r$ and $b_s$ as the same edge. This operation is particularly important as we will prove that if a graph is planar, then so is any graph or \gls{multigraph} obtained by contracting an \gls{edge}. The proof for this will be provided.
%TODO put refrence to proof for contraction of planar graph is planar

\subsubsection{Terminology: subgraph}
We can say that a graph $A$ is a subgraph of another graph $B$, if all \glspl{vertex} in $A$ are also in $B$ and each \glspl{edge} in $A$ is also in $B$.

\subsubsection{Groundwork}
Here we would like to prove some lemmas that we would like to use while proving Kuratowski's theorem.

First we would like to add a single limit point to

\printnoidxglossaries

>>>>

\documentclass{article}
\input{../../definitions}
\input{../topology}
\input{glossary}
\title{Topological Graph Theory}
\begin{document}
\maketitle
\tableofcontents

@@

\section{Overview}
\subsection{Graphs}
Before we talk about topological graph theory with any level of understanding we must first understand what a \gls{graph} is.

A \gls{graph} is generally defined as a \gls{set} of \glspl{vertex} combined with a \gls{set} of \glspl{edge} between \glspl{vertex}, however here it may be more useful to think about them visually with a simple representation.

@@

Once we have drawn this we have a representation of a \gls{graph}. If we were to move the \glspl{vertex} around on the paper but leave them having the same \glspl{edge} (the same \glspl{vertex} are connected to the point as they were before). we would be left with the same \gls{graph}. That is to say, it doesn't mater where we put a \gls{vertex} on our sheet, the \gls{graph} exists independently of the representation we draw.

@@

I first came across this question in the 8\textsuperscript{th} grade, and it was presented to us during geometry class. While undoubtedly an interesting problem, it is quite misleading to try and think of this as a geometry problem. Instead we will try and reduce it to a \gls{graph} problem.

Let us start by thinking of every island as a \gls{vertex} and every bridge as an \gls{edge}. We find the following \gls{graph}.

@@

It is worth noting two things about the above diagram. First that the labels on the \glspl{vertex} and \glspl{edge} are unrelated to the problem, but have been added simply to make referring to parts of the \gls{graph} much easier. Second that whatever the above image depicts, does not fit our definition of a \gls{graph}.

Notice that \gls{edge} $2$ and \gls{edge} $3$ both connect \gls{vertex} $A$ to $B$, as well \glspl{edge} $5$ and $6$ do for $B$ and $C$. This is a strict violation of our definition for a \gls{graph}. The issue of course then comes to what would one call such a beast as this where, presumably, one is able to have as many connections between any pair of \glspl{vertex} and could even have connections from a \gls{vertex} to itself.

I am particularly glad that you're paying enough attention to notice that the diagram does not depict a \gls{graph}. This is what we will refer to as a \gls{multigraph}. It is worth noting that \glspl{graph} are a type of \gls{multigraph}, so anything we show to be true for all \glspl{multigraph}, is also true for all \glspl{graph}.

@@

Consider the following \gls{graph}.

@@

This \gls{graph} is called a \gls{complete} as every \gls{vertex} is connected to every other \gls{vertex} by an \gls{edge}; in fact this \gls{graph} in particular is called $K_4$ as it is the complete four vertex \gls{graph}. We would like to find out if we can draw this above \gls{graph} without having any lines crossing. We can in fact draw this \gls{graph} without any intersections and for any skeptics who may being reading this, the below is $K_4$ without any \glspl{edge} intersecting.

@@

So if this \gls{graph} can be drawn without intersection, can any \gls{graph} be drawn without intersections? If some can and some can't how do we tell which can be drawn and which can not?

\section{Kuratowski's Theorem}
Kuratowski's theorem tell us exactly what \glspl{graph} can be drawn on a plane or sphere. We say either a plane or sphere as it turns out to be the same question. %TODO prove that drawing on plane and sphere is same question
Now before we tackle the theorem let's lay down some terminology.

\subsection{Terminology}

@@

We say this \gls{graph} has two \glspl{face}. One \gls{face} is easy to see and contained within triangle $ABC$. The other \gls{face} is a bit harder to see, however it is all the space outside of the triangle $ABC$. We may call this second face the outer \gls{face} for this drawing of the \gls{graph}, however it is also possible to draw it such that the inner face may be the outer face. You can think of this as kind of turning the triangle inside out.

A face only makes sense to talk about for a \gls{graph} that can be embedded in a certain space. For example the \gls{graph} $K_4$ may look like it has $5$ faces in the following drawing.

@@

	\path[every node/.style={font=\sffamily\small}]
	(1) edge node {} (4)
	edge node {} (3)
	(2) edge node {} (1)
	edge node {} (4)
	(3) edge node {} (2)
	(4) edge node {} (3);
	\end{tikzpicture}
\end{center}
This would be wrong as a face assumes a drawing without \glspl{edge} intersecting. This can be rectified if the \gls{graph} can be drawn without intersection and thus we find this drawing.
\begin{center}
	\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
	thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]

@@

	\path[every node/.style={font=\sffamily\small}]
	(1) edge node {} (4)
	edge [bend right = 100]node {} (3)
	(2) edge node {} (1)
	edge node {} (4)
	(3) edge node {} (2)
	(4) edge node {} (3);
	\end{tikzpicture}
\end{center}
This means there are actually $4$ \glspl{face} in $K_4$.

@@

Let us consider the following \gls{graph}.

@@

This can then be reduced into a \gls{graph} again by simply treating $b_r$ and $b_s$ as the same edge. This operation is particularly important as we will prove that if a \gls{graph} is planar, then so is any \gls{graph} or \gls{multigraph} obtained by contracting an \gls{edge}. The proof for this will be provided.
%TODO put refrence to proof for contraction of planar graph is planar

\subsubsection{Terminology: subgraph}
We can say that a \gls{graph} $A$ is a \gls{subgraph} of another \gls{graph} $B$, if all \glspl{vertex} in $A$ are also in $B$ and each \glspl{edge} in $A$ is also in $B$.

\subsection{Groundwork}
Here we would like to prove some lemmas that we would like to use while proving Kuratowski's theorem.

\begin{theorem}[Spherical and Planar graphs are the same] \label{sphere=plane}
	If and only if a \gls{graph} may be drawn on a plane with no \glspl{edge} intersection then it may be drawn on a sphere without self intersection.
\end{theorem}

\begin{proof}
	First we know from topology that the punctured sphere is homeomorphic to the plane, planarity of a graph is the same as saying that the graph may be drawn on a punctured sphere without any self intersection. Now it's trivial to say that anything drawn on a punctured sphere may be drawn on the sphere. Now it will suffice to show only that a graph drawing will never cover the entire sphere as then there exists some at which to puncture.
\end{proof}

\printnoidxglossary
